---
title: "ICS_Cytokine_Analysis"
author: "Taryn McLaughlin"
date: "9/21/2018"
output: 
  html_document: default
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE, message=FALSE, warning=FALSE, comment=NA)
library(dplyr)
library(knitr)
library(ggplot2)
library(gridExtra)
library(GGally)
library(ggpubr)
library(cowplot)
library(tidyr)
library(reshape2)
```

```{r source}
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/tryna function.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/pvals_TB-SM.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/plot_functions.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/comp_hc_ltbi.R")
```

```{r load data, echo=FALSE}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean/")

CD4<-read.csv("ICS_CD4_Cytokine_clean.csv")

CD8<-read.csv("ICS_CD8_Cytokine_clean.csv")

MFI<-read.csv("ICS_CD4_MFI_clean.csv")

Compass<-read.csv("ICS_Compass_scores_complete_clean.csv")

GD<-read.csv("ICS_GD_Cytokine_clean.csv")

data<-read.csv("ICS_DonorInfo_clean.csv")
data$TB<-gsub("BCG|Unknown", "Na誰ve", data$TB)
data$TB<-factor(data$TB, levels=c("Na誰ve","Healthy Control","LTBI","Active TB"))
data$SM<-gsub("US|Not US|Unknown", "Na誰ve", data$SM)
data$SM<-factor(data$SM, levels=c("Na誰ve","Negative","Schisto+"))
Donor<-data
```

##PMA & Ionomycin
```{r}
library(knitr)
#creates a list of character strings for each of the columns of interest
colnames<-names(dplyr::select(CD4, -X, -Sample, -Donor, -Stim, -TB, -SM))
for(col in colnames){#this will loop through each character string in the list
    a<-subset(CD4, CD4$Stim=="PMA") #subsets the data to only contain the unstimulated samples
    #inserts each character string into the column argument of the pvals function written above
    b<-pvals_6(a, col)
    if(dim(b)[1]!=0){
    print(kable(b, caption=col))#will return the output for each character string in the loop
    y<-plot_with_all_stats(a, col)+labs(title=col)
    print(y)}
    }
```


##Whole Cell Lysate
```{r}
library(knitr)
#creates a list of character strings for each of the columns of interest
colnames<-names(dplyr::select(CD4, -X, -Sample, -Donor, -Stim, -TB, -SM))
for(col in colnames){#this will loop through each character string in the list
    a<-subset(CD4, CD4$Stim=="WCL") #subsets the data to only contain the unstimulated samples
    #inserts each character string into the column argument of the pvals function written above
    b<-pvals_6(a, col)
    if(dim(b)[1]!=0){
    print(kable(b, caption=col))#will return the output for each character string in the loop
    y<-plot_with_all_stats(a, col)+labs(title=col)
    print(y)}
    }
```

##CFP-10 & ESAT-6
```{r}
library(knitr)
#creates a list of character strings for each of the columns of interest
colnames<-names(dplyr::select(CD4, -X, -Sample, -Donor, -Stim, -TB, -SM))
for(col in colnames){#this will loop through each character string in the list
    a<-subset(CD4, CD4$Stim=="PEP") #subsets the data to only contain the unstimulated samples
    #inserts each character string into the column argument of the pvals function written above
    b<-pvals_6(a, col)
    if(dim(b)[1]!=0){
    print(kable(b, caption=col))#will return the output for each character string in the loop
    y<-plot_with_all_stats(a, col)+labs(title=col)
    print(y)}
    }
```

```{r}
df<-mutate_TB(CD4)
grid.arrange(
    plot_2(subset(df, df$Stim=="PMA"), "disease", "IFNg")+
        labs(y="IFNg+ CD4+ T cells (%)", title="PMA & Ionomycin"),
    plot_2(subset(df, df$Stim=="PEP"), "disease", "IFNg")+
        labs(y="", title="CFP-10 & ESAT-6"),
    ncol=2) 

grid.arrange(
    plot_2(subset(df, df$Stim=="PMA"), "disease", "G_T")+
        labs(y="IFNg+TNFa+ CD4+ T cells (%)", title="PMA & Ionomycin"),
    plot_2(subset(df, df$Stim=="PEP"), "disease", "G_T")+
        labs(y="", title="CFP-10 & ESAT-6"),
    ncol=2) 

grid.arrange(
    plot_2(subset(df, df$Stim=="PMA"), "disease", "x_4_13_T")+
        labs(y="IL4+TNFa+ CD4+ T cells (%)", title="PMA & Ionomycin"),
    plot_2(subset(df, df$Stim=="PEP"), "disease", "x_4_13_T")+
        labs(y="", title="CFP-10 & ESAT-6"),
    ncol=2) 
```


#Extra Graphs

##Compass Graphs
```{r}
#CD4 Mtb Responses as measured by Compass
DF<-dplyr::filter(Compass, TB != "N" & Stim == "WCL" & celltype == "CD4")
my_comparisons <- list(c("LTBI SM", "LTBI X"), c("TB SM", "TB X"),
                       c("TB SM", "LTBI SM"), c("TB X", "LTBI X"))

p1<-plot_compass(DF, "FS")+
    labs(y="Functionality Score")

p2<-plot_compass(DF, "PFS")+
    labs(y="Polyfunctionality Score")

grid.arrange(p1, p2, ncol=2, top="Whole Cell Lysate Scores")

DF<-dplyr::filter(Compass, TB != "N" & Stim == "PEP" & celltype == "CD4")

p1<-plot_compass(DF, "FS")+
    labs(y="Functionality Score")

p2<-plot_compass(DF,"PFS")+
    labs(y="Polyfunctionality Score")

grid.arrange(p1, p2, ncol=2, top="Peptide Pool Scores")
```

##Boolean Graphs
The first graph is just meant to give a snapshot of all the data at once to see if there are weird anomalous data points that we will have to examine. I did not include all the individual data points on any of the graphs because it made it look incredibly cluttered, however I did allow outliers to display.  

```{r Boolean Graphs, cache=TRUE, echo=FALSE, warning=FALSE, fig.width=20}
pma<-filter(CD4, Stim=="PMA")
wcl<-filter(CD4, Stim=="WCL")
pep<-filter(CD4, Stim=="PEP")
triple_boolean_plot(pma)
triple_boolean_plot(wcl)
triple_boolean_plot(pep)
```


##T cell Lineage Specific Grouped Analysis
These are my standard lineage specific groups. 
```{r Grouped TH Graphs, echo=FALSE}
library(dplyr)
library(reshape2)
library(ggplot2)
library(knitr)
colnames<-c("Cyt.Freq", "TH1.Freq", "TH1.2.Freq", "TH2.Freq")
### PMA
for(col in colnames){#this will loop through each character string in the list
    #inserts each character string into the column argument of the pvals function written above
    name<-gsub("\\.", "+", col)
    name<-gsub("Freq","", name)
    y<-plot_with_all_stats(subset(CD4, CD4$Stim=="PMA"), col)+
        labs(title=paste("Frequency of ", name, " CD4+ T cells"), 
             subtitle="PMA & Ionomycin",
             x="",
             y=paste(name, " CD4+ T cells (%)"))
    print(y) #will return the output for each character string in the loop
}
##### WCL
for(col in colnames){#this will loop through each character string in the list
    #inserts each character string into the column argument of the pvals function written above
    name<-gsub("\\.", "+", col)
    name<-gsub("Freq","", name)
    y<-plot_with_all_stats(subset(CD4, CD4$Stim=="WCL"), col)+
        labs(title=paste("Frequency of ", name, " CD4+ T cells"), 
             subtitle="WCL",
             x="",
             y=paste(name, " CD4+ T cells (%)"))
    print(y) #will return the output for each character string in the loop
}
##### Peptide
for(col in colnames){#this will loop through each character string in the list
    #inserts each character string into the column argument of the pvals function written above
    name<-gsub("\\.", "+", col)
    name<-gsub("Freq","", name)
    y<-plot_with_all_stats(subset(CD4, CD4$Stim=="PEP"), col)+
        labs(title=paste("Frequency of ", name, " CD4+ T cells"), 
             subtitle="PEP",
             x="",
             y=paste(name, " CD4+ T cells (%)"))
    print(y) #will return the output for each character string in the loop
}
```
\newpage

##Individual Cytokines
These are the frequencies of cells producing individual cytokines without considering whether or not they are single positive, double positive etc. So any cell producing IL-4 is included in the IL-4 analysis.

```{r Individual Cytokines but not single producers}
library(dplyr)
library(reshape2)
library(ggplot2)
library(knitr)
colnames<-c("IFNg", "TNFa", "IL4", "IL13")

### PMA
for(col in colnames){#this will loop through each character string in the list
    #inserts each character string into the column argument of the pvals function written above
    y<-plot_with_all_stats(subset(CD4, CD4$Stim=="PMA"), col)+
        labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 
             subtitle="PMA & Ionomycin",
             x="",
             y=paste(col, "+ CD4+ T cells (%)"))
    print(y) #will return the output for each character string in the loop
}
####WCL
for(col in colnames){#this will loop through each character string in the list
    #inserts each character string into the column argument of the pvals function written above
    name<-gsub("\\.", "+", col)
    name<-gsub("Freq","", name)
    y<-plot_with_all_stats(subset(CD4, CD4$Stim=="WCL"), col)+
        labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 
             subtitle="WCL",
             x="",
             y=paste(col, "+ CD4+ T cells (%)"))
    print(y) #will return the output for each character string in the loop
}

### PEP
for(col in colnames){#this will loop through each character string in the list
    #inserts each character string into the column argument of the pvals function written above
    name<-gsub("\\.", "+", col)
    name<-gsub("Freq","", name)
    y<-plot_with_all_stats(subset(CD4, CD4$Stim=="PEP"), col)+
        labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 
             subtitle="PEP",
             x="",
             y=paste(col, "+ CD4+ T cells (%)"))
    print(y) #will return the output for each character string in the loop
}
```


```{r Single and Double Producers Lin Spec graphs, echo=FALSE}
library(dplyr)
library(reshape2)
library(ggplot2)
library(knitr)

colnames<-c("G_T" , "IL4_IL13" )

### PMA
for(col in colnames){#this will loop through each character string in the list
    #inserts each character string into the column argument of the pvals function written above
    name<-gsub("_", "+", col)
    y<-plot_with_all_stats(subset(CD4, CD4$Stim=="PMA"), col)+
        labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 
             title="PMA & Ionomycin",
             x="",
             y=paste(col, "+ CD4+ T cells (%)"))
    print(y) #will return the output for each character string in the loop
}
####WCL
for(col in colnames){#this will loop through each character string in the list
    #inserts each character string into the column argument of the pvals function written above
    name<-gsub("_", "+", col)
    y<-plot_with_all_stats(subset(CD4, CD4$Stim=="WCL"), col)+
        labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 
             subtitle="WCL",
             x="",
             y=paste(col, "+ CD4+ T cells (%)"))
    print(y) #will return the output for each character string in the loop
}

### PEP
for(col in colnames){#this will loop through each character string in the list
    #inserts each character string into the column argument of the pvals function written above
    name<-gsub("_", "+", col)
    y<-plot_with_all_stats(subset(CD4, CD4$Stim=="PEP"), col)+
        labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 
             subtitle="PEP",
             x="",
             y=paste(col, "+ CD4+ T cells (%)"))
    print(y) #will return the output for each character string in the loop
}
```
